---
title: "DataSci 306, Homework 2"
author: "Max Han, maxhan"
output:
  pdf_document: default
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(nycflights13)
library(babynames)
```


## Question 1 (2 points)

For this challenge we will use the `babynames` dataset that is already loaded in the above code block (refer: `library(babynames)`)

```{r}
babynames |> head(100)
```
To learn more on this data, run `?babynames` from your R console

(a) Plot the frequency (the `n` variable in the data, denotes the frequency) of the name `Kamala` over the years. (1 point)

```{r}
babynames |> filter(name == "Kamala") |>
  ggplot(aes(x = year, y =n)) + 
  geom_line(color = "black", fill = "white") +
  labs(title = "Graph for frequency of name 'Kamala' over time",
       subtitle = "by Max Han",
       x = "Year",
       y = "Frequency")
```


(b) Find the three most popular baby names in the year 1964 (1 point)

```{r}
babynames |> filter(year == 1964) |> slice_max(n, n = 3)
```


## Question 2 (5 points)

Manipulating `flights` data

In this problem, we will explore `nycflights` data.  This data is loaded for you in the above code block (refer: `library(nycflights13)`)

```{r}
flights |> head()
```

Learn more on this data by invoking `?flights` on your console. Then answer the below questions



(a) How many flights had dep_delay (i.e., departure delay) value of more than 2 hours? (0.5 point)


```{r}
flights |> filter(dep_delay > 120) |> count()
```

(b) Mutating a new column (1.5 pt)

Add a new column called `status` that marks the flight as 'Delayed' if the dep_delay value is more than 10. If the dep_delay is < 0, then mark it is as 'Early'. Rest of them mark it as 'On-time'

Then select only `dep_delay` and the new column `status` to display the first 30 records

**Hint:** Will you use `if_else` or `ifelse`?  There are other ways to solve it as well (you could solve it differently too) but we have so far understood the use of if-else statements.


```{r}

flights_1 <- mutate(flights, 
                    status = if_else(dep_delay > 10, "Delayed", 
                                     if_else(dep_delay < 0, "Early", "On-time")))
select(flights_1, dep_delay, status) |> head(30)
```


(c) Delays by origin (2 points)

Here is a plot showing the departure delays over time, attempting to breakdown the data by origin using color

```{r}
flights |>
  ggplot(aes(x = month, y = dep_delay, color = origin)) +
  geom_point() + 
  theme(legend.position="bottom") + 
  labs(y = "Departure Delay",
        x = "Month")
```

What would be a better way to display this data? Implement a better plot and explain why it is better.
Also answer, which airport shows the biggest issues with departure delays? 

Hint: Explore how `facet` chart can help. Also for over-plotting situation, which geometry would work better? What about the color? Can you set a different variable for color to make it more readable?

```{r}
flights |>
  ggplot(aes(x = month, y = dep_delay, color = origin)) +
  geom_point() +
  facet_wrap(~ origin) + 
  theme(legend.position="bottom") + 
  labs(y = "Departure Delay",
        x = "Month")
```


(d) More filtering (1 point)

Select only those rows that belong to carriers 'UA', 'AA' and 'B6' and show the distribution of flights for these three carriers using a bar chart

```{r}
flights |> filter(carrier %in% c("UA", "AA", "B6")) |>
  ggplot(aes (x = carrier, fill = carrier)) +
  geom_bar()
```


## Question 3 - Challenge problem (3 points)
**Spooky flights**

Define a flight to be spooky if it was in transit at 13:13 h (i.e., 1:13 pm) on Friday the 13th of any month.
You can assume the flight is in transit between its `dep_time` and `arr_time`. 

Hint: You need to extract the weekday. Explore functions from `lubridate` package to extract the weekday.

```{r}
# If year, month, and day are given, checks if the date is Friday.
is_friday <- function(year, month, day) {
  date <- make_date(year, month, day)
  return (wday(date, label = TRUE) == "Fri")
}

# Checks if the range between dep_time and arr_time contains 1313.
is_spooky_time <- function(dep_time, arr_time) {
  return (dep_time < 1313 & arr_time > 1313)
}

# Filters the data according to the given criteria.
spooky_flights <- flights |> filter(day == 13 & 
                                    is_friday(year, month, day) & 
                                    is_spooky_time(dep_time, arr_time))

spooky_flights
```


