---
title: "DataSci 306, Homework 5"
author: "Max Han, maxhan"
output:
  pdf_document: default
---

```{r setup, include=T}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(lubridate)
library(nycflights13)
```

**Reshaping and Joins**

In almost all problems it would be easier for you if you reshape your dataframe using pivot_wider and/or pivot_longer and/or use joins


## Problem 1 (3 points)

(a) Let us revisit the storms data and let us view the total number of storms that are of category 1 through 5 across each available month as shown in the fig-1.png enclosed with this HW (1 point)
 
```{r}
head(storms)
```
 
 
```{r}
storms |> filter(category %in% c(1:5)) |> group_by(category, month) |> summarise(n = n(), .groups = "drop") |> pivot_wider(names_from = month, values_from = n) |> mutate(across(everything(), ~ coalesce(.x, 0)))


```

See fig-1.png that is part of this HW



```{r}
storms |> filter(category %in% c(1:5)) |> group_by(category, month) |> summarise(n = n(), .groups = "drop") |> pivot_wider(names_from = month, values_from = n) |> mutate(across(where(is.numeric), ~ coalesce(.x, 0)))
```

(b) Iris dataset (2 points)
For this problem, we will work on the iris data. To learn more on iris, please read the documentation with `?iris` as always

```{r}
iris |> head()
```


Add columns to show the ratio of Length/Width for both sepal and petal for the iris dataset such that the sepal_ratio shows the ratio of Sepal.Length/Sepal.Width and the petal_ratio shows the ratio of Petal.Length/Petal.Width. Then reshape the dataset to make it longer with only three columns; Species, ratio and value; where ratio column contains both petal_ratio and sepal_ratio and then reproduce the following plot. (2 points)

See ratio.png that is part of this HW

```{r}
iris |> mutate(sepal_ratio = Sepal.Length / Sepal.Width, petal_ratio = Petal.Length / Petal.Width) |> 
  pivot_longer(cols = c(sepal_ratio, petal_ratio), names_to = "ratio", values_to = "value") |>
  select(Species, ratio, value) |> ggplot(aes(y = value, x = Species, color = Species)) +
  geom_boxplot() + facet_wrap(~ratio)
```


## Problem 2 (4 points)

(a) Age of aircraft vs Cancellations (2 points)

We will look into the `flights` dataset again for this problem

Is there a relation between age of the aircraft and proportion of cancellations? Answer this question using a suitable chart

Hint: This data was collected in 2013. Derive the age of the aircraft using the year of manufacture of the aircraft and then find the proportion of flights cancelled.

```{r}
flights |> print()

```
```{r}
flights_age <- flights |>
  left_join(planes, by = "tailnum") |>
  filter(!is.na(year.y)) |>
  mutate(age = 2013 - as.numeric(year.y), .before = year.x) 

flights_cancellation <- flights_age %>%
  group_by(age) %>%
  summarise(total_flights = n(),
            cancellations = sum(is.na(dep_time)),  
            prop_cancelled = cancellations / total_flights)

ggplot(flights_cancellation, aes(x = age, y = prop_cancelled)) +
  geom_point() +
  geom_line() +
  labs(title = "Proportion of Cancellations vs. Aircraft Age",
       x = "Aircraft Age (Years)",
       y = "Proportion of Flights Cancelled") +
  theme_minimal()
flights_cancellation
```
The graphs shows no pattern. Thus, we can conclude that there is no relationship between plane age and the proportion of cancellation.



(b) Dep Delay due to weather? (2 points)

Filter all the flights that have a departure delay of more than 60 minutes, and then find if weather parameters are related to delays. The weather parameters to consider are humid, wind_speed, wind_gust, precip, pressure, visib.  Using a facet_grid, and a suitable geom, slice and dice the dataset based on the three origins and the given weather factors and explain which of these weather parameters influence delays if any.

Hint: You may want to join along with reshape to get the end result.


```{r}
delayed_flights_weather <- flights |> filter(dep_delay > 60) |> left_join(weather, by = c("origin", "year", "month", "day", "hour"))

ggplot(delayed_flights_weather, aes(x = wind_speed, y = dep_delay)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "blue") +
  facet_grid(origin ~ .) +
  labs(title = "Departure Delay vs Wind Speed",
       x = "Wind Speed (mph)",
       y = "Departure Delay (minutes)")

ggplot(delayed_flights_weather, aes(x = humid, y = dep_delay)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "blue") +
  facet_grid(origin ~ .) +
  labs(title = "Departure Delay vs Humidity",
       x = "Humidity",
       y = "Departure Delay (minutes)")

ggplot(delayed_flights_weather, aes(x = wind_gust, y = dep_delay)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "blue") +
  facet_grid(origin ~ .) +
  labs(title = "Departure Delay vs Wind Gust",
       x = "Wind Gust (mph)",
       y = "Departure Delay (minutes)")

ggplot(delayed_flights_weather, aes(x = precip, y = dep_delay)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "blue") +
  facet_grid(origin ~ .) +
  labs(title = "Departure Delay vs Precipitation",
       x = "Precipitation (inches)",
       y = "Departure Delay (minutes)")

ggplot(delayed_flights_weather, aes(x = pressure, y = dep_delay)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "blue") +
  facet_grid(origin ~ .) +
  labs(title = "Departure Delay vs Pressure",
       x = "Pressure (millibars)",
       y = "Departure Delay (minutes)")

ggplot(delayed_flights_weather, aes(x = visib, y = dep_delay)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(method = "lm", color = "blue") +
  facet_grid(origin ~ .) +
  labs(title = "Departure Delay vs Visibility",
       x = "Visibility (miles)",
       y = "Departure Delay (minutes)")
```


All the plots shows flat line for linear relationship, which means that slopes are 0. Thus, we can conclude weather doesn't affect delay time.



## Problem 3 (3 points)

In this problem, we will explore the `relig_income` dataset. As always learn more on this dataset by running `?relig_income`

```{r}
relig_income |> head()
```

When you look at the dataset, you will notice that this data is not tidy. First make the data tidy by creating new columns called 'income_range' and 'count' to rearrange the existing column names.

Once that step is done, filter out the category 'Don’t know/refused' category that is present in both religion and income_range and then get all the religions that have a total count across all income groups greater or equal to 200 and show the distribution of these religions across the various income groups using a suitable plot. Show a distribution chart that makes it simple to compare religious distribution across income levels.

```{r}
joined <- relig_income |> pivot_longer(-religion, names_to = "income_range", values_to = "count")
joined
```

```{r}
cleaned <- joined |> filter(income_range != "Don't know/refused" & count != "Don't know/refused")
cleaned
```

```{r}
over_200_count_religion <- cleaned |> group_by(religion) |> summarise(count_over_200 = sum(count, na.rm = T)) |> filter(count_over_200 >= 200) |> select(religion)
over_200_count_religion
```

```{r}
filtered_income <- cleaned |> inner_join(over_200_count_religion, by = "religion")
filtered_income
```


```{r}
filtered_income |> ggplot(aes(x = income_range, y = count, fill = religion)) +
  geom_bar(stat = "identity", position = "dodge") +
  labs(title = "Distribution of Religions Across Income Levels",
       x = "Income Range",
       y = "Count",
       fill = "Religion")
```

