---
title: "DATASCI 306, Fall 2024, Final Group Project"
author: "3 "
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dplyr)
library(gridExtra)
library(tidyverse)
library(lubridate)
library(reshape2)
```

Throughout this course, you've dedicated yourself to refining your analytical abilities using R programming language. These skills are highly coveted in today's job market!  

Now, for the semester project, you'll apply your learning to craft a compelling `Data Story` that can enrich your portfolio and impress prospective employers. Collaborating with a team (up to 5 members of your choosing), you'll construct a Data Story akin to the example provided here: <https://ourworldindata.org/un-population-2024-revision>

Data is already in the `data` folder. This data is downloaded from: <https://population.un.org/wpp/Download/Standard/MostUsed/>

You'll conduct Exploratory Data Analysis (EDA) on the provided data. The provided article already includes 6 diagrams.  Show either the line or the map option for these 6 charts. You may ignore the table view. I'm also interested in seeing how each team will expand upon the initial analysis and generate additional 12 insightful charts that includes US and any other region or country that the author did not show.  For e.g., one question you may want to answer is; US population is expected to increase to 421 million by 2100. You may want to show how the fertility rate and migration may be contributing to this increase in population.

**Deliverable**

**1. Requirement-1 (2 pt) **
Import the data given in the .xlxs file into two separate dataframes;

* one dataframe to show data from the `Estimates` tab 
* one dataframe to show data from the `Medium variant` tab 

Hint: Some of the steps you may take while importing include:

* skip the first several comment lines in the spread sheet
* Importing the data as text first and then converting the relevant columns to different datatypes in step 2 below.

``` {r}
file_path <- "data/WPP2024_GEN_F01_DEMOGRAPHIC_INDICATORS_COMPACT.xlsx"

library(readxl)

data_estimates <- read_excel(file_path, sheet = "Estimates", col_names = TRUE, skip = 16, col_types = "text")
data_medium_variant <- read_excel(file_path, sheet = "Medium variant", col_names = TRUE, skip = 16, col_types = "text")

```


**2. Requirement-2 (5 pt)**

You should show at least 5 steps you adopt to clean and/or transform the data. Your cleaning should include:

* Renaming column names to make it more readable; removing space, making it lowercase or completely giving a different short name; all are acceptable.
* Removing rows that are irrelevant; look at rows that have Type value as 'Label/Separator'; are those rows required?
* Removing columns that are redundant; For e.g., variant column
* Converting text values to numeric on the columns that need this transformation

You could also remove the countries/regions that you are not interested in exploring in this step and re-save a smaller file in the same `data` folder, with a different name so that working with it becomes easier going forward.

Explain your reasoning for each clean up step.

```{r}
#remove "Label/Separator"
data_estimates <- data_estimates %>% filter(Type != "Label/Separator")
data_medium_variant <- data_medium_variant %>% filter(Type != "Label/Separator")

# Remove the Notes Column because data has no value and appears to either be NA or random characters or numbers with no significance
data_estimates <- data_estimates %>% select(-Notes)
data_medium_variant <- data_medium_variant %>% select(-Notes)

#Removes the variant column because it is redundant because the data is already split by sheet
data_estimates <- data_estimates %>% select(-Variant)
data_medium_variant <- data_medium_variant %>% select(-Variant)
```

```{r}
#Removes the "Total Population, as of 1 January (thousands)" column because it is redundant because is is the only column with a January date and every other column is dated to July
data_estimates <- data_estimates %>% select(-'Total Population, as of 1 January (thousands)')
data_medium_variant <- data_medium_variant %>% select(-'Total Population, as of 1 January (thousands)')
```

```{r}
# Convert all columns after column 10 to numeric for `data_estimates` because they are all numbers
data_estimates <- data_estimates %>%
  mutate(across(.cols = 9:ncol(.), as.numeric))

# Convert all columns after column 10 to numeric for `data_medium_variant` because they are all numbers
data_medium_variant <- data_medium_variant %>%
  mutate(across(.cols = 9:ncol(.), as.numeric))

# View the head of the cleaned data
head(data_estimates)
head(data_medium_variant)
```

```{r}
#All columns dates are July, so the July in the column name is redundant
data_estimates <- data_estimates %>%
  rename_with(~ str_replace_all(., ", as of 1 July ", " "), .cols = 0:62)
data_medium_variant <- data_medium_variant %>%
  rename_with(~ str_replace_all(., ", as of 1 July ", " "), .cols = 0:62)
```


**3.  Requirement-3 (3 pt)**
Replicate the 6 diagrams shown in the article.  Show only the '2024' projection values where ever you have both '2022' and '2024' displayed. 
Show only the diagrams that are shown on the webpage with default options. 

```{r}
population24_world <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` == "World") |> 
  select(Year,'Total Population (thousands)') |> rename(TotalPopulation = `Total Population (thousands)`)

population24_africa <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` == "Africa") |> 
  select(Year,'Total Population (thousands)') |> rename(TotalPopulation = `Total Population (thousands)`)

population24_asia <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` == "Asia") |> 
  select(Year,'Total Population (thousands)') |> rename(TotalPopulation = `Total Population (thousands)`)

population24_europe <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` == "Europe") |> 
  select(Year,'Total Population (thousands)') |> rename(TotalPopulation = `Total Population (thousands)`)

population24_northernAmerica <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` == "Northern America") |> 
  select(Year,'Total Population (thousands)') |> rename(TotalPopulation = `Total Population (thousands)`)

population24_latin <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` == "Latin America and the Caribbean") |> 
  select(Year,'Total Population (thousands)') |> rename(TotalPopulation = `Total Population (thousands)`)

```


####### Diagram 1 ########

```{r fig.width=10, fig.height=5}
plot_world <- ggplot(data = population24_world, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(8*10^6, 10^7, by = 0.5 * 10 ^ 6), labels = paste0(seq(8, 10, by = 0.5), " billion")) +
  labs(title = "The medium population projection from the UN's World Population Prospects in its 2024 publication",
       x = "Year",
       y = "Total Population")
print(plot_world)

```


####### Diagram 2 ########

```{r}
plot_world <- ggplot(data = population24_world, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(8*10^6, 10^7, by = 0.5 * 10 ^ 6), labels = paste0(seq(8, 10, by = 0.5), "B")) +
  labs(title = "World")
```


```{r}
plot_africa <- ggplot(data = population24_africa, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(1.5*10^6, 3.5*10^6, by = 0.5 * 10 ^ 6), labels = paste0(seq(1.5, 3.5, by = 0.5), "B")) +
  labs(title = "Africa (UN)")
```


```{r}
plot_asia <- ggplot(data = population24_asia, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(4.7*10^6, 5.3*10^6, by = 0.1 * 10 ^ 6), labels = paste0(seq(4.7, 5.3, by = 0.1), "B")) +
  labs(title = "Asia (UN)")
```


```{r}
plot_europe <- ggplot(data = population24_europe, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(6*10^5, 7*10^5, by = 0.5 * 10 ^ 5), labels = paste0(seq(600, 700, by = 50), "M")) +
  labs(title = "Europe (UN)")
```


```{r}
plot_northAmerica <- ggplot(data = population24_northernAmerica, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(4*10^5, 4.6*10^5, by = 0.2 * 10 ^ 5), labels = paste0(seq(400, 460, by = 20), "M")) +
  labs(title = "North America (UN)")
```


```{r}
plot_latin <- ggplot(data = population24_latin, aes(x = Year, y = TotalPopulation)) +
  geom_point( color = "blue") +
  geom_line(color = "blue") +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  scale_y_continuous(breaks = seq(6.2*10^5, 7.4*10^5, by = 0.2 * 10 ^ 5), labels = paste0(seq(620, 740, by = 20), "B")) +
  labs(title = "Latin America and the Caribbean (UN)")
```


```{r fig.width=18, fig.height=10}
grid.arrange(plot_world, plot_africa, plot_asia, plot_europe, plot_northAmerica, plot_latin, ncol = 3, nrow = 3)
```
####### Diagram 3 ########

```{r}
birthRate <- rbind(data_estimates, data_medium_variant) |> filter(Year >= 1950)
regions <- c("Africa", "World", "Asia", "Latin America and the Caribbean", "Northern America", "Europe")
combinedData3 <- birthRate |> filter(`Region, subregion, country or area *` %in% regions)

```



```{r fig.width=12, fig.height=5}

ggplot(data = combinedData3, aes(x = Year, y = `Total Fertility Rate (live births per woman)`, color = `Region, subregion, country or area *`)) +
  geom_line() +
  scale_x_continuous(breaks = c(1950, seq(1980, 2100, by = 20))) +
  labs(
    title = "Fertility Rate: Children per Woman (1950 to 2100)",
    x = "Year",
    y = "Total Fertility Rate",
    color = "Region"
  ) +
  theme_minimal()
```

####### Diagram 4 ########

```{r}
regions <- c("China", "India")
combinedData4 <- birthRate |> filter(`Region, subregion, country or area *` %in% regions)
```


```{r fig.width=12, fig.height=5}
ggplot(data = combinedData4, aes(x = Year, y = `Total Population (thousands)`, color = `Region, subregion, country or area *`)) +
  geom_point(size = 0.5) +
  geom_line() +
  scale_x_continuous(breaks = c(1950, seq(1980, 2100, by = 20))) +
  scale_y_continuous(breaks = seq(2*10^5, 1.6*10^6, by = 2 * 10 ^ 5), labels = c(paste0(seq(200, 800, by = 200), " million"), 
                                                                                 paste0(seq(1, 1.6, by = 0.2), " billion"))) +
  labs(
    title = "Population, 1950 to 2100",
    x = "Year",
    y = "Population",
    color = "Country"
  ) +
  theme_minimal()
```

####### Diagram 5 ########

```{r}
birthRate <- rbind(data_estimates, data_medium_variant) |> filter(Year >= 1950, Year <= 2023)
regions <- c("Northern America", "Europe", "Latin America and the Caribbean", "Asia", "World", "Africa")
combinedData5 <- birthRate |> filter(`Region, subregion, country or area *` %in% regions)

```


```{r fig.width=12, fig.height=5}
ggplot(data = combinedData5, aes(x = Year, y = `Life Expectancy at Birth, both sexes (years)`, color = `Region, subregion, country or area *`)) +
  geom_point(size = 0.5) +
  geom_line() +
  scale_x_continuous(breaks = seq(1950, 2023, by = 10)) +
  scale_y_continuous(breaks = seq(0, 70, by = 10), labels = paste0(seq(0, 70, by = 10)), " years") +
  labs(
    title = "Population, 1950 to 2100",
    x = "Year",
    y = "Population",
    color = "Country"
  ) +
  theme_minimal()
```



####### Diagram 6 ########

```{r}
total <- rbind(data_estimates, data_medium_variant) |> filter(Year >= 1950, Year <= 2023)
gibraltar <- total |> filter(`Region, subregion, country or area *` == "Gibraltar")
low_income <- total |> filter(`Region, subregion, country or area *` == "Low-income countries" )
combinedData6 <- rbind(low_income, gibraltar)

```


```{r fig.width=12, fig.height=5}
ggplot(data = combinedData6, aes(x = Year, y = `Net Number of Migrants (thousands)`, color = `Region, subregion, country or area *`)) +
  geom_point(size = 0.5) +
  geom_line() +
  scale_x_continuous(breaks = seq(1950, 2024, by = 10)) +
  scale_y_continuous(breaks = seq(-2 * 10 ^ 3, 5 * 10 ^ 2, by = 5 * 10 ^ 2), 
                     labels = c("-2 million", "-1.5 million", "-1 million", "-500,000", "0", "500,000")) +
  labs(
    title = "Anual net migration, 1950 to 2023",
    x = "Year",
    y = "Population",
    color = "Region"
  ) +
  theme_minimal()
```
####### Diagram 6 (REDO) ########

```{r}
data_estimates |> filter(`Region, subregion, country or area *` == "Ukraine") |> 
  mutate(Annual_Net_Migration = `Net Number of Migrants (thousands)` * 1000) |> 
  ggplot(aes(x = Year, y = Annual_Net_Migration)) + 
  geom_point() +
  geom_line() + 
  scale_y_continuous(
    breaks = c(0, -1000000, -2000000, -3000000, -4000000, -5000000),         
    labels = c("0", "-1 Million", "-2 Million", "-3 Million", "-4 Million", "-5 Million")  
  ) + 
  scale_x_continuous(
    limits = c(1950, 2023),
    breaks = c(1950, 1960, 1970, 1980, 1990, 2000, 2010, 2023)
  ) +
  labs(
    title = "Anual net migration, 1950 to 2023",
    x = "",
    y = "",
  ) +
  theme_minimal()
```


**4.  Requirement-4 (12 pt)**

Select United States related data, and any other country or region(s) of your choosing to perform EDA. Chart at least 12 additional diagrams that may show relationships like correlations, frequencies, trend charts, between various variables with plots of at least 3 different types (line, heatmap, pie, etc.). Every plot should have a title and the x/y axis should have legible labels without any label overlaps for full credit. 

Summarize your interpretations after each chart.

############### The Goal of EDA ################

This analysis delves into how the population and population key factors in the United States, China, India, and Russia would change from 2024 to 2100.

```{r}
regions <- c("United States of America" , "China", "India", "Russian Federation")
data_eda <- data_medium_variant |> filter(Year >= 2024, `Region, subregion, country or area *` %in% regions) |> 
  rename(TotalPopulation = `Total Population (thousands)`, Region = `Region, subregion, country or area *`)

numeric_columns <- data_eda |> select(is.numeric) 

```


```{r fig.width=12, fig.height=5}
ratio2024 <- data_eda |> filter(Year == 2024) |>
  ggplot(aes(x = "", y = TotalPopulation, fill = Region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round(TotalPopulation * 100 / sum(TotalPopulation), 1), "%")), 
            position = position_stack(vjust = 0.5), size = 4) +
  labs(title = "Population Ratio over Region before Pandemic (2024)") +
  theme_void()

ratio2100 <- data_eda |> filter(Year == 2100) |>
  ggplot(aes(x = "", y = TotalPopulation, fill = Region)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar(theta = "y") +
  geom_text(aes(label = paste0(round(TotalPopulation * 100 / sum(TotalPopulation), 1), "%")), 
            position = position_stack(vjust = 0.5), size = 4) +
  labs(title = "Pie Chart for Population Ratio over Region after Pandemic (2100)") +
  theme_void()

grid.arrange(ratio2024, ratio2100, nrow = 2)
```

The charts above compare the population ratios of the four countries in 2024 and 2100, highlighting significant shifts. The most notable trends are the rapid increase in India's population ratio and the steep decline in China's. India's population ratio surges from 43.2% in 2024 to 56% in 2100, reflecting sustained growth and demographic momentum. Conversely, China's population ratio experiences a dramatic drop, decreasing from 42.2% to 23.6%, likely driven by declining birth rates and an aging population. These changes underscore the contrasting population trajectories of the two nations over the observed period.

```{r fig.width=12, fig.height=5}
bar2024 <- data_eda |> filter(Year == 2024) |> group_by(Region) |> ggplot(aes(y = TotalPopulation, x = Region, fill = Region)) +
  scale_y_continuous(breaks = seq(0, 2 * 10^6, by = 0.25 * 10 ^ 6), labels = paste0(seq(0, 2, by = 0.25), "B")) +
  geom_bar(stat = "identity") +
  labs(title =  "Population Comparison Over Region in 2024")

bar2100 <- data_eda |> filter(Year == 2100) |> group_by(Region) |> ggplot(aes(y = TotalPopulation, x = Region, fill = Region)) +
  scale_y_continuous(breaks = seq(0, 2 * 10^6, by = 0.25 * 10 ^ 6), labels = paste0(seq(0, 2, by = 0.25), "B")) +
  geom_bar(stat = "identity") +
  labs(title =  "Population Comparison Over Region in 2100")

grid.arrange(bar2024, bar2100, nrow = 2)

```
The bar charts above illustrate the population figures for each country in 2024 and 2100, providing insights into population shifts over time. A striking observation is the sharp population decline in China, with its population reducing to less than half by 2100. In contrast, the other countries—India, the United States, and Russia—show relatively minor changes in their populations over the same period, reflecting more stable demographic trends. This dramatic reduction in China's population highlights the profound impact of its low birth rates and aging population.

```{r fig.width=12, fig.height=5}
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = TotalPopulation, color = Region)) +
  geom_line() +
  geom_point() +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Population Comparison Over Region",
       x = "Year", 
       y = "Population (in thousands)")

```
The graph above depicts the population trends for each country over time. India demonstrates a notable increase in population, peaking around 2065, before gradually declining to return to its 2024 population level by 2100. In contrast, China experiences a continuous and sharp population decline throughout the period. Meanwhile, the United States and Russia maintain relatively stable populations, showing minimal variation over time.

```{r}
data_eda |> filter(Year %in% c(2024, 2100)) |> group_by(Region, Year) |> 
  select(Region, Year, TotalPopulation) |> head(10)
```


```{r fig.width=20, fig.height=20}

correlation_matrix <- cor(numeric_columns, use = "complete.obs")
correlation_melted <- melt(correlation_matrix)

ggplot(data = correlation_melted, aes(x = Var1, y = Var2, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "steelblue", high = "blue", mid = "white", 
                       midpoint = 0, limit = c(-1, 1), space = "Lab", 
                       name = "Correlation") +
  theme_minimal() + 
  labs(title = "Correlation Heatmap for USA", x = "Attributes", y = "Attributes") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1))
```

From the above correlation table, population growth rate, total death, fertility rate, migration rate, and sex ratio has a strong relation with population change in our analysis context. The accompanying graphs illustrate the dynamics of population growth rates, offering a clear visualization of these shifts across countries.

```{r fig.width=12, fig.height=5} 
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = `Population Growth Rate (percentage)`, color = Region)) +
  geom_line() +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Population Growth Rate Comparison Over Region",
       x = "Year",
       Y = "Growth Rate in Percentage")
```


For China, the population growth rate shows a steep decline over time, transitioning into negative territory from 2024 and continuing to drop further into the century. India initially exhibits the highest population growth rate compared to other regions but experiences a steady decline throughout the timeline. Growth rate becomes negative from 2060. Russia maintains a negative rate but relatively stable population growth rate throughout the observed period, increasing slightly territory after 2070. The United States shows a moderately positive and stable growth rate over the entire period, with only a slight downward trend toward the end of the century.


```{r fig.width=12, fig.height=5}
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = `Total Deaths (thousands)` * 100 / TotalPopulation, color = Region)) +
  geom_line() +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Total Death Rate Comparision Over Region",
       x = "Year",
       y = "Death Rate in Percentage")
```

The graphs illustrate distinct trends in death rates across the regions. Russia exhibits the highest death rate in 2024, but this begins to decline steadily from 2045 onward. The United States shows an increasing death rate up to 2050, after which it stabilizes at a relatively consistent level. In contrast, China, despite starting with a relatively low death rate, experiences a rapid upward trajectory over time, reflecting a significant long-term increase. India stands out as having the lowest death rate up to 2065; however, it also shows a gradual rise as the timeline progresses, indicating the influence of demographic changes and an aging population. 

```{r fig.width=12, fig.height=5}
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = `Total Fertility Rate (live births per woman)`, color = Region)) +
  geom_line() +
  
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Fertility Rate Comparison Over Region",
       x = "Year",
       y = "Fertility Rate (live birth per women)")
```

India demonstrates the highest birth rate among the analyzed countries, though it experiences a steady decline over time. The United States follows with the second-highest birth rate, maintaining a relatively stable trend throughout the observed period. Russia ranks third, also showing a consistent and stable birth rate over time. In contrast, China holds the lowest birth rate since 2024 but exhibits a noticeable upward trend during the observed period, signaling a relative increase in fertility rates.

```{r fig.width=12, fig.height=5}
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = `Net Migration Rate (per 1,000 population)`, color = Region)) +
  geom_line() +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Net Migration Rate Comparison Over Region",
       x = "Year", 
       y = "Net Migration Rate (per 1,000 population)")
```
A notable observation in this graph is the significant shift in Russia's emigration rate. Around 2025, Russia's rate sharply increases from -1.7 to 2.3 and remains relatively stable at approximately 2.0 through 2100. In contrast, the other countries do not exhibit any substantial changes in their emigration rates over the observed period, maintaining relatively consistent trends. This sharp rise in Russia's emigration rate stands out as a distinct anomaly compared to the stability seen in other regions.

```{r fig.width=12, fig.height=5}
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = `Median Age (years)`, color = Region)) +
  geom_line() +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Median Age Comparison Over Region",
       x = "Year", 
       y = "Age")
```
A striking feature in the graph is the rapid increase in median age observed in China and India. In China, the median age rises sharply from 40 to 60, reflecting significant aging within the population. India, starting with the lowest median age of 35 in 2024, experiences a notable increase, reaching 45 and becoming the second highest among the analyzed countries. In contrast, the United States and Russia exhibit relatively stable trends, maintaining median ages between 40 and 45 throughout the observed period. These patterns highlight the demographic shifts driven by aging populations and declining fertility rates in China and India, compared to more stable trends in the United States and Russia.

```{r fig.width=12, fig.height=5}
data_eda |> group_by(Region) |> ggplot(aes(x = Year, y = `Population Sex Ratio (males per 100 females)`, color = Region)) +
  geom_line() +
  scale_x_continuous(breaks = c(2024, seq(2020, 2100, by = 10))) +
  labs(title =  "Sex Ratio Comparison Over Region",
       x = "Year", 
       y = "Sex Ratio (males per 100 females)")
```

The graph above highlights a striking trend in Russia's sex ratio, which rises significantly from 87 males per 100 females to 95 males per 100 females over the observed period. In contrast, the sex ratios for the other countries remain relatively stable, fluctuating within a narrow range of 100 males per 100 females to 105 males per 100 females. This notable increase in Russia's sex ratio suggests a shift in its demographic composition, while the stability in other countries indicates balanced gender distributions over time.

############### Miscellaneous EDA ################

```{r}
data_estimates |> 
  filter(`Region, subregion, country or area *` == c("More developed regions", "Less developed regions")) |> 
  mutate(development = if_else(`Region, subregion, country or area *` == "More developed regions", "More Developed Region", "Less Developed Region")) |> 
  ggplot(aes(x = Year, y = `Infant Mortality Rate (infant deaths per 1,000 live births)`, color = development)) +
  geom_line() +
  scale_y_continuous(breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100, 110, 120, 130, 140)) +
  labs(title = "Infant Mortality Rates Over Time Grouped By Regional Development Status",
    color = "Development"
    ) 
```
This chart illustrates a steady decline in infant mortality rates in both more developed and less developed regions over the past seven decades. This trend is likely attributed to advancements in infant healthcare and medical technology over this time period. Additionally, the significantly lower infant mortality rate in more developed regions can be explained by their greater access to medical developments and advanced healthcare infrastructure than in less developed regions. The steeper decline in the infant mortality rate for less developed regions suggests a larger relative improvement in medical care. This is likely driven by the accelerated adoption of modern medical practices and technologies in less developed regions, which had a larger area of potential improvement compared to the already advanced healthcare infrastructure in more developed regions. 

```{r}
data_estimates |> 
  filter(`Region, subregion, country or area *` %in% c("Western Europe", "Eastern Europe"), Year >= 1987, Year <= 1996) |> 
  mutate(region = if_else(`Region, subregion, country or area *` == "Western Europe", "Western Europe", "Eastern Europe"), 
         grouping = if_else(Year <= 1991, "5 Years Before", "5 Years After"),
         grouping = factor(grouping, levels = c("5 Years Before", "5 Years After"))) |> 
  group_by(region, grouping) |> 
  summarize(mean_net_migration_per_1000 = mean(`Net Migration Rate (per 1,000 population)`, na.rm = T)) |>
  ggplot(aes(x = grouping, y = mean_net_migration_per_1000, fill = region)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~ region) +
  theme_minimal() +
  labs(x = "Time Period in Relation to the Fall of the USSR",
       y = "Average Net Migration Per 1,000 People",
       title = "Migration Changes in Eastern and Western Europe Around the Fall of the USSR",
       fill = "Region"
    ) 
```
In Eastern Europe, net migration was slightly positive in the period before the USSR's collapse. This is likely a result of the challenging conditions within the USSR before its collapse, which contributed to limited migration. After the fall of the USSR, net migration increased in Eastern Europe, suggesting a growing desire to migrate to newly independent countries that were no longer under Soviet control. In contrast, Western Europe has a significantly higher net migration in both periods, which is likely a result of the abundant economic opportunities that were available to migrants coming to Western Europe. However, the decrease in net migration to Western Europe following the USSR's collapse, coupled with the increase in Eastern Europe's net migration, further supports the notion that migrants were attracted to newly independent Eastern European countries during this period of transition.

```{r}
data_estimates |> 
  filter(`Region, subregion, country or area *` %in% c("High-income countries", "Low-income countries")) |> 
  mutate(income = if_else(`Region, subregion, country or area *` == "High-income countries", "High Income Country", "Low Income Country")) |>
  group_by(income, Year) |> 
  summarize("Difference in Female and Male Life Expectancy" = `Female Life Expectancy at Birth (years)` - `Male Life Expectancy at Birth (years)`) |>
  ggplot(aes(x = Year, y = `Difference in Female and Male Life Expectancy`, color = income)) +
  geom_line() +
  scale_y_continuous(limit = c(0, 7)) +
  labs(color = "Income",
       title = "Advantage In Female Life Expectancy Based On National Income"
    ) +
  theme_minimal()
```
This chart demonstrates that in both high and low income countries women consistently outlive men on average. In high income countries, the female live expecatnacy advantage over men stabilizes around 1955 and steadily rises till its peak of about 7 years in 1980 where it stays till the early 90s. It then gradually declines to a five year advantage in the modern day. This trend may reflect changing societal factors, including health and lifestyle behaviors, as well as a decreasing gender gap in workplace stress and mortality risks. In contrast, the female life expectancy advantage in low income countries remains relatively steady at around 4 years from its stabilization in 1955 till around the 2010. After 2010, there is a significant rise as the difference increases to around 5 years, approximately matching that of high income countries in the modern day. This convergence between high and low income countries may be a result of improvements in healthcare access and reduced maternal mortality rates in low income regions. 

```{r}
Countries <- c("United States of America", "France","Canada", "India", "Pakistan","Kenya")

Death_rate_data <- data_estimates %>%
  filter(Year <= 2023) %>%
  bind_rows(data_medium_variant %>% filter(Year > 2023)) %>%
  filter(`Region, subregion, country or area *` %in% Countries) %>%
  select(`Region, subregion, country or area *`, Year, `Crude Death Rate (deaths per 1,000 population)`) %>%
  rename(
    Country = `Region, subregion, country or area *`,
    Death_rate = `Crude Death Rate (deaths per 1,000 population)`
  ) %>%
  filter(!is.na(Death_rate)) %>%
  ggplot(aes(x = Year, y = Death_rate , color = Country)) +
  geom_line() +
  scale_x_continuous(
    breaks = c(1950, 1980, 2000, 2020, 2040, 2060, 2080, 2100),
    limits = c(1950, 2100)
  ) +
  scale_y_continuous(
    limits = c(0, 40), 
    breaks = seq(0, 40, 10)
  ) +
  labs(
    title = "Death Rate (1950-2100)",
    subtitle = "Comparison of Developed and Developing Countries",
    x = "Year",
    y = "Death Rate (deaths per 1,000 population)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 8),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )
print(Death_rate_data)
```
The graph shows the trends in death from 1950 to 2100 for six countries. We can see the differences between developed countries (United States of America, France and Canada) and developing countries (Pakistan, India, Kenya). The developing countries generally had much higher death rate than developed countries due to poor health infrastructure and environment but in the present day it has gone down and the death rate is similar to developed countries' death rate. But the death rate for both sections of countries is predicted to increase in the future and developing nations' death rate is going to increase more than developed nations' death rate.

```{r}
Countries <- c("United States of America", "Spain","Canada", "Bangladesh", "Pakistan","India")

Fertility_rate_data <- data_estimates %>%
  filter(Year <= 2023) %>%
  bind_rows(data_medium_variant %>% filter(Year > 2023)) %>%
  filter(`Region, subregion, country or area *` %in% Countries) %>%
  select(`Region, subregion, country or area *`, Year, `Total Fertility Rate (live births per woman)`) %>%
  rename(
    Country = `Region, subregion, country or area *`, 
    Fertility_rate = `Total Fertility Rate (live births per woman)`
  ) %>%
  filter(!is.na(Fertility_rate)) %>%
  ggplot(aes(x = Year, y = Fertility_rate , color = Country)) +
  geom_line() +
  scale_x_continuous(
    breaks = c(1950, 1980, 2000, 2020, 2040, 2060, 2080, 2100),
    limits = c(1950, 2100)
  ) +
  scale_y_continuous(
    limits = c(0, 10), 
    breaks = seq(0, 10, 5)
  ) +
  labs(
    title = "Fertility Rate (1950-2100)",
    subtitle = "Comparison of Developed and Developing Countries",
    x = "Year",
    y = "Fertility Rate (live births per woman)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 8),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )
print(Fertility_rate_data)

```
The graph shows the trends in fertility rates from 1950 to 2100 for six countries. We can see the differences between developed countries (United States of America, Spain and Canada) and developing countries (Pakistan, Bangladesh, India). Developing countries generally had higher fertility rate and has gone down in the present time but Pakistan still has the highest rate and will remain so.

```{r}
Countries <- c("United States of America", "China","Canada", "Nigeria", "Indonesia","India")

Infant_deaths_data <- data_estimates %>%
  filter(Year <= 2023) %>%
  bind_rows(data_medium_variant %>% filter(Year > 2023)) %>%
  filter(`Region, subregion, country or area *` %in% Countries) %>%
  select(`Region, subregion, country or area *`, Year, `Infant Deaths, under age 1 (thousands)`) %>%
  rename(
    Country = `Region, subregion, country or area *`, 
    Infant_deaths = `Infant Deaths, under age 1 (thousands)`
  ) %>%
  filter(!is.na(Infant_deaths)) %>%
  ggplot(aes(x = Year, y = Infant_deaths , color = Country)) +
  geom_line() +
  scale_x_continuous(
    breaks = c(1950, 1980, 2000, 2020, 2040, 2060, 2080, 2100),
    limits = c(1950, 2100)
  ) +
  scale_y_continuous(
    limits = c(0, 5000), 
    breaks = seq(0, 5000, 1000)
  ) +
  labs(
    title = "Infant Deaths (1950-2100)",
    subtitle = "Comparison of Developed and Developing Countries",
    x = "Year",
    y = "Infant Deaths, under age 1 (thousands)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 8),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )
print(Infant_deaths_data)
```
The graph shows the trends in infant deaths from 1950 to 2100 for six countries. We can see the differences between developed countries (United States of America, China and Canada) and developing countries (Nigeria, Indonesia and India). China and India had the largest amount of infant deaths because of huge populations but it has gone down and will further go do down and become similar to other countries in the future.

```{r}
india_data <- data_estimates %>%
  filter(`Region, subregion, country or area *` == "India") %>%
  select(Year, 
         `Total Fertility Rate (live births per woman)`, 
         `Life Expectancy at Birth, both sexes (years)`, 
         `Population Growth Rate (percentage)`) %>%
  rename(
    FertilityRate = `Total Fertility Rate (live births per woman)`,
    LifeExpectancy = `Life Expectancy at Birth, both sexes (years)`,
    PopulationGrowthRate = `Population Growth Rate (percentage)`
  ) %>%
  filter(!is.na(FertilityRate) & !is.na(LifeExpectancy) & !is.na(PopulationGrowthRate))

cor_data <- india_data %>%
  select(-Year) %>%
  cor(use = "complete.obs") %>%
  as.data.frame() %>%
  rownames_to_column(var = "Variable1") %>%
  gather(Variable2, Correlation, -Variable1)

ggplot(cor_data, aes(x = Variable1, y = Variable2, fill = Correlation)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0, 
                       name = "Correlation") +
  labs(
    title = "Correlation Heatmap: India Life Expectancy, Fertility, and Population Growth",
    x = "",
    y = ""
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    plot.title = element_text(hjust = 0.5, size = 9, face = "bold")
  )
```
The correlation between life expectancy and fertility rate appears to be strongly negative, which aligns with many demographic trends in developing countries like India, where improved life expectancy often coincides with improved health, education, and access to family planning which lower fertility rates. There is a strong correlation between population growth and fertility rate, which is logical since more births means an increase in population. The correlation between life expectancy and pop growth rate is interestingly weakly positive(purple), which seems to suggest that life expectancy could lead to increase in population growth, but is not always the case. Other factors like decreasing fertility rate may balance the population growth over time like in developed countries such as Japan. 

```{r}
countries <- c("United States of America", "Democratic Republic of the Congo", "Japan")

combined_growth_data <- data_estimates %>%
  filter(Year <= 2023) %>%
  bind_rows(data_medium_variant %>% filter(Year > 2023)) %>%
  filter(`Region, subregion, country or area *` %in% countries) %>%
  select(`Region, subregion, country or area *`, Year, `Population Growth Rate (percentage)`) %>%
  rename(
    Country = `Region, subregion, country or area *`,
    GrowthRate = `Population Growth Rate (percentage)`
  ) %>%
  filter(!is.na(GrowthRate))

ggplot(combined_growth_data, aes(x = Year, y = GrowthRate, color = Country)) +
  geom_line(size = 1.2) +
  scale_x_continuous(
    breaks = c(1950, 1980, 2000, 2020, 2040, 2060, 2080, 2100), 
    limits = c(1950, 2100)
    ) +
  scale_y_continuous(limits = c(-1, 6), breaks = seq(-1, 6, 1)) +
  labs(
    title = "Population Growth Rate (1950-2100)",
    subtitle = "Comparing the USA, DR Congo, and Japan",
    x = "Year",
    y = "Population Growth Rate (%)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 7, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 10),
    axis.text = element_text(size = 10),
    axis.title = element_text(size = 12),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )
```
The DRC displays the highest growth rates, peaking in the late 20th century due to high fertility rates, though it gradually declines over time. Japan's growth rate steadily decreases, turning negative in recent decades, reflecting an aging population, low fertility, and population decline. Meanwhile, the United States maintains a relatively stable growth rate, slowly declining but remaining positive, driven by moderate fertility and immigration. These trends highlight the DRC's rapid population expansion due to being a developing country, Japan's demographic challenges, and the USA's balanced growth.

```{r}
countries <- c("United States of America", "Germany", "Democratic Republic of the Congo", "Japan", "India", "Nigeria", "China", "Cuba")

age_data <- data_estimates %>%
  filter(Year <= 2023) %>%
  bind_rows(data_medium_variant %>% filter(Year > 2023)) %>%
  filter(`Region, subregion, country or area *` %in% countries) %>%
  select(`Region, subregion, country or area *`, Year, `Median Age (years)`) %>%
  rename(
    Country = `Region, subregion, country or area *`,
    MedianAge = `Median Age (years)`
  ) %>%
  filter(!is.na(MedianAge))

ggplot(age_data, aes(x = Year, y = MedianAge, color = Country)) +
  geom_line(size = 1.2) +
  scale_x_continuous(
    breaks = c(1950, 1980, 2000, 2020, 2040, 2060, 2080, 2100),
    limits = c(1950, 2100)
  ) +
  scale_y_continuous(
    limits = c(10, 60), 
    breaks = seq(10, 60, 10)
  ) +
  labs(
    title = "Median Age Over Time (1950-2100)",
    subtitle = "Comparison of Developed and Developing Countries",
    x = "Year",
    y = "Median Age (Years)",
    color = "Country"
  ) +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, size = 12, face = "bold"),
    plot.subtitle = element_text(hjust = 0.5, size = 8),
    axis.text = element_text(size = 8),
    axis.title = element_text(size = 10),
    legend.title = element_text(size = 8),
    legend.text = element_text(size = 8)
  )
```
The graph shows the trends in median age from 1950 to 2100 for six countries, highlighting differences between developed (United States, Germany, Japan) and developing nations (Democratic Republic of the Congo, India, Nigeria). Developed countries generally exhibit higher median ages throughout, reflecting aging populations due to lower fertility rates and higher life expectancy. Conversely, developing nations start with significantly lower median ages, indicative of younger populations with higher birth rates, but experience gradual increases as demographic transitions occur.

The small dip in the United States' median age around 1970 corresponds to the Baby Boom, a post-World War II event, marked by a significant rise in birth rates. This influx of younger individuals temporarily reduced the median age during this period. Over time, as the Baby Boomers aged and birth rates declined, the median age began rising again. 

**5. Requirement-5 (2 pt)**
Having developed a strong understanding of your data, you'll now create a machine learning (ML) model to predict a specific metric. This involves selecting the most relevant variables from your dataset.

The UN's World Population Prospects provides a range of projected scenarios of population change. These rely on different assumptions in fertility, mortality and/or migration patterns to explore different demographic futures. Check this link for more info: https://population.un.org/wpp/DefinitionOfProjectionScenarios

You can choose to predict the same metric the UN provides (e.g., future population using fertility, mortality, and migration data). Compare your model's predictions to the UN's. 

How significantly do your population projections diverge from those of the United Nations? Provide a comparison of the two.  If you choose a different projection for which there is no UN data to compare with, then this comparison is not required.

```{r}
library(caret)
set.seed(123)

target <- "Total Population (thousands)"
predictors <- c("Total Population (thousands)",
                "Total Fertility Rate (live births per woman)",
                "Infant Mortality Rate (infant deaths per 1,000 live births)",
                "Under-Five Mortality (deaths under age 5 per 1,000 live births)",
                "Mortality before Age 40, both sexes (deaths under age 40 per 1,000 live births)",
                "Male Mortality before Age 40 (deaths under age 40 per 1,000 male live births)",
                "Female Mortality before Age 40 (deaths under age 40 per 1,000 female live births)",
                "Mortality before Age 60, both sexes (deaths under age 60 per 1,000 live births)",
                "Male Mortality before Age 60 (deaths under age 60 per 1,000 male live births)",
                "Female Mortality before Age 60 (deaths under age 60 per 1,000 female live births)",
                "Mortality between Age 15 and 50, both sexes (deaths under age 50 per 1,000 alive at age 15)",
                "Male Mortality between Age 15 and 50 (deaths under age 50 per 1,000 males alive at age 15)",
                "Female Mortality between Age 15 and 50 (deaths under age 50 per 1,000 females alive at age 15)",
                "Mortality between Age 15 and 60, both sexes (deaths under age 60 per 1,000 alive at age 15)",
                "Male Mortality between Age 15 and 60 (deaths under age 60 per 1,000 males alive at age 15)",
                "Female Mortality between Age 15 and 60 (deaths under age 60 per 1,000 females alive at age 15)",
                "Net Number of Migrants (thousands)",
                "Net Migration Rate (per 1,000 population)")

data <- data_estimates[predictors]
train_index <- createDataPartition(data$`Total Population (thousands)`, p = 0.7, list = F)
train_data <- data[train_index, ]
test_data <- data[-train_index, ]


```


```{r}
model <- train(
  `Total Population (thousands)` ~ ., 
  data = test_data,
  method = "rf",
  trControl = trainControl(method = "cv", number = 5)
)  

```


```{r}
saveRDS(model, file = "trained_model.rds")
print(model)
summary(model$finalModel)
```

```{r}
predictions <- predict(model, newdata = test_data)
head(predictions)
```

```{r}
actuals <- test_data$`Total Population (thousands)`

rmse <- sqrt(mean((predictions - actuals)^2))  
mae <- mean(abs(predictions - actuals))       
r_squared <- cor(predictions, actuals)^2    

cat("RMSE:", rmse, "\n")
cat("MAE:", mae, "\n")
cat("R-squared:", r_squared, "\n")

```


```{r}
compare <- data.frame(actual = actuals, prediction = predictions)
ggplot(data = compare, aes(x = actual, y = predictions, color = "blue"))+
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "red")

```

**6.  Requirement-5 (1 pt)**

**Conclusion**

Your analysis should conclude with a summary of key findings. I'm especially interested in any novel insights you uncover that go beyond the article's original conclusions.

Here's key findings we get through EDA.

The analysis reveals significant demographic shifts across the four countries by 2100. China's population is projected to decline sharply to 0.625 billion, driven by a high death rate and persistently low fertility rates. Russia's migration rate is expected to surge dramatically from -1.75 persons per 1,000 population to 2.25 persons per 1,000 population by 2025, a level that remains stable through 2100. The median age in China will rise notably to 60 years, while the United States, India, and Russia maintain median ages around 40 years, reflecting relatively stable aging trends. Additionally, Russia's sex ratio will undergo a significant shift, increasing from 85 males per 100 females to 95 males per 100 females by 2100, marking a major demographic transformation.

Furthermore, the analysis of infant mortality rates across countries in various states of development demonstrate that modernization of healthcare especially in less developed countries contributes to declines in infant mortality rates, which can contribute to an increase in overall population projections. Additionally, the migration changes seen in Eastern and Western Europe as a result of the collapse of the USSR demonstrates the positive trends in net migration that are a result of dismantling failing political regimes. When examining the number of years women outlive men in high income versus low income countries, it can be argued that, due to a combination of narrowing gender disparities in high income nations and significant improvements in healthcare access and maternal mortality reduction in low-income countries, there has been a convergence in recent years in the life expectancy gap between women and men across these economic contexts.

The analysis for death rate shows that developing countries had much higher rates but have gone down in the present day and both sections of countries’ rates will increase in the future. Additionally, the results for fertility rate shows that developing countries had much higher rates but they’ve also gone down and will become similar to developed countries’ rates. Furthermore, analysis for infant deaths shows that countries with more population had higher rates especially in India and China but they’ve gone down as those countries have become more developed and are predicted to become similar to other countries.

The analysis highlights key demographic trends and offers insights into potential future shifts across countries. The strong negative correlation between life expectancy and fertility rate suggests that as health, education, and access to family planning continue to improve in developing nations like India, fertility rates are likely to decline further, leading to slower population growth over time. Although population growth is positively correlated with fertility rates, the weakly positive correlation between life expectancy and growth indicates that in developed countries like Japan, an aging population and decreasing fertility rates has already led to a decline in population growth. The United States is projected to approach a 0.5% growth rate by 2040 and Japan's growth rate to continue to remain negative the next 80 years. In contrast, countries like the DRC are expected to continue experiencing rapid population expansion, driven by high fertility rates, though this growth may gradually stabilize as the country progresses through its demographic transition.

Projections for median age trends reveal that developed countries such as Japan, Germany, and the United States will see continued increases in median age, reaching well past 40 by 2040, reflecting aging populations and low fertility rates. Developing countries like Nigeria and the DRC, while currently characterized by younger populations, are anticipated to see steady increases in median age as fertility rates decline and life expectancy improves. These increases will continue from 2020 onwards till 2100, reaching an expected median age of 35, still well below projected median age for already developed countries by that time. Historical trends, such as the U.S. Baby Boom, demonstrate how temporary shifts in fertility can alter demographic profiles, and similar patterns may emerge in other regions. Looking ahead, the aging populations in developed nations and the youth-driven growth in developing regions will shape global demographic dynamics, potentially impacting economic development, healthcare systems, and social structures.


**7. Extra Credit (1 pt)**
Develop an interactive Shiny app to visualize your machine learning model's projections. The app must include at least one interactive widget (e.g., dropdown, radio buttons, text input) allowing users to select a variable value (such as country/region) and view the corresponding projections.


**Submission**

* You will upload the zip file containing finals.Rmd file and its PDF as a deliverable to Canvas. If you created a shiny app for predictions, you will add those files also to your zip file.

* You will present your findings by creating a video of a maximum 15 minutes duration, explaining the code and the workings of your project; all team members should explain their part in the project to receive credit. You will share the URL of the video on Canvas for us to evaluate. An ideal way to create this video would be to start a Zoom meeting, start recording, and then every member share their screen and explain their contribution.


It is not necessary to prepare slides (if you do it doesn't hurt) for the presentation. You may speak by showing the diagrams and/or code from your Posit project.  Every team member should explain their part in the project along with the insights they derived by explaining the charts and summaries for full credit to each member.

Your project will be evaluated for clean code, meaningful/insightful EDA and predictions.

**Note:**

* Each plot must be accompanied by a summary that clarifies the rationale behind its creation and what insights the plot unveils. Every diagram should possess standalone significance, revealing something more compelling than the other charts
* After the deadline, instructors will select the top three outstanding analytics projects. The teams responsible for these exceptional analyses will have their video shared with the class

**We will not accept submissions after the deadline; December 10th 4 pm**



